---
title: "Posture Project"
author: "Giovani Gutierrez"
date: "2023-02-15"
output: 
  github_document:
    toc: true
    df_print: "kable"
    keep_html: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      tidy = TRUE,
                      fig.align = 'center')

indent1 = '    '
indent2 = paste(rep(indent1, 2), collapse='')
indent3 = paste(rep(indent1, 3), collapse='')
```

# Getting Started

## Setup
```{r libraries}
library(tidyverse)
library(tidymodels)
library(readxl)
library(janitor)
library(visdat)
library(naniar)
library(corrplot)
library(patchwork)
library(glmnet)
library(kknn)
library(rpart)
library(rpart.plot)
library(ranger)
library(vip)
tidymodels_prefer()
theme_set(theme_bw())

set.seed(123) # set seed
```

## Import Data
```{r read-in data}
data1 <- 
  read_xls('C:/Users/giova/Desktop/PSTAT 131/predicting_personality/Data/Posture_Data.xls') %>% 
  clean_names() # read in raw data from .xls file

data1 %>% 
  write.csv(file = 'C:/Users/giova/Desktop/PSTAT 131/predicting_personality/Data/raw_data.csv') # write raw data to .csv file

head(data1) # preview data
```

## Tidy Data & Missing Values
```{r missing values, fig.width = 10}
data1 <- 
  data1 %>% 
  replace_with_na_all(condition = ~.x == 'X') # replace 'X' character values with NA values

data1 %>% 
  vis_miss() # visualize missing values

data1 <- 
  data1 %>% 
  select(age:posture & -mbti) %>% 
  drop_na() # drop missing and irrelevant variables

data1 <- 
  data1 %>% 
  mutate(sex = factor(sex)) %>% 
  mutate(activity_level = ordered(activity_level,
                                  levels = c('Low', 'Moderate', 'High'))) %>% 
  mutate_at(c('pain_1', 'pain_2', 'pain_3', 'pain_4'), 
            as.numeric) %>% 
  mutate(e_i = factor(ifelse(e > i, 'Extrovert', 'Introvert'))) %>% 
  mutate_at(c('s', 'n', 't', 'f', 'j', 'p'), as.numeric) %>% 
  mutate(posture = factor(posture)) %>% 
  select(-e & -i) # coerce variables into appropriate type

data1 %>% 
  vis_dat() # visualize variable types

data1 %>% 
  write.csv(file = 'C:/Users/giova/Desktop/PSTAT 131/predicting_personality/Data/clean_data.csv') # write clean data to .csv file

head(data1) # preview cleaned data
```

* `age`: Age of participant (in years)
* `height`: Height of participant (in inches)
* `weight`: Weight of participant (in pounds)
* `sex`: Sex of the participant (male or female)
* `activity_level`: Amount of daily physical activity performed by a participant (low, moderate, or high)
* `pain_1`: Pain in the neck reported on a scale from 0 (low) to 10 (high) by a participant
* `pain_2`: Pain in the thoracic reported on a scale from 0 (low) to 10 (high) by a participant
* `pain_3`: Pain in the lumbar reported on a scale from 0 (low) to 10 (high) by the participant
* `pain_4`: Pain in the sacrum reported on a scale from 0 (low) to 10 (high) by the participant
* `s`: A participant's sensing characteristic on a scale from 1 (low) to 26 (high)
* `i`: A participant's intuition characteristic on a scale from 1 (low) to 26 (high)
* `t`: A participant's thinking characteristic on a scale from 1 (low) to 26 (high)
* `f`: A participant's feeling characteristic on a scale from 1 (low) to 26 (high)
* `j`: A participant's judging characteristic on a scale from 1 (low) to 26 (high)
* `p`: A participant's perceiving characteristic on a scale from 1 (low) to 26 (high)
* `posture`: Type of posture exhibited by a participant (ideal posture (A), kyphosis-lordosis (B), flat-back (C), and sway-back (D))
* `e_i`: Categorization of participant as an extrovert or introvert; based on whichever score was higher on their MBTI questionnaire


# Exploratory Data Analysis (EDA)

## Correlation Between Features

```{r correlation plot}
tmwr_cols <- colorRampPalette(c("#91CBD765", "#CA225E"))

data1 %>% 
  select_if(is.numeric) %>% 
  cor() %>% 
  corrplot(col = tmwr_cols(200), 
           tl.col = "black",
           method = 'circle', 
           type = 'lower',
           diag = FALSE)
```

## Visual EDA
```{r visual eda}
ggplot(data = data1, aes(fill = e_i, x = e_i)) + 
  geom_bar(colour = 'black') +
  scale_fill_brewer(palette = "Pastel1") +
  guides(fill = FALSE) +
  labs(x = NULL, y = 'Number of Participants')

ggplot(data = data1, aes(fill = sex, x = e_i)) + 
  geom_bar(position = "fill", colour = 'black') +
  scale_fill_brewer(palette = "Pastel1") + 
  labs(x = NULL, y = 'Proportion of Participants', fill = 'Sex')

ggplot(data = data1, aes(fill = posture, x = e_i)) + 
  geom_bar(position = "fill", colour = 'black') +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = NULL, y = 'Proportion of Participants', fill = 'Posture Type')

ggplot(data = data1, aes(fill = activity_level, x = posture)) + 
  geom_bar(position = 'fill', colour = 'black') +
  labs(x = 'Posture Type', fill = 'Activity Level') +
  facet_grid(. ~ e_i) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = 'Proportion of Participants', fill = 'Activity Level')

ggplot(data = data.frame(data1, transmute(data1, s_n = factor(ifelse(s > n, 'sensing', 'inuitive')))), aes(x = e_i, fill = s_n)) + 
  geom_bar(position = 'fill', colour = 'black') +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = '', y = 'Proportion of Participants', fill = NULL)

ggplot(data = data.frame(data1, transmute(data1, t_f = factor(ifelse(t > f, 'thinking', 'feeling')))), aes(x = e_i, fill = t_f)) + 
  geom_bar(position = 'fill', colour = 'black')  +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = NULL, y = 'Proportion of Participants', fill = NULL)

ggplot(data = data.frame(data1, transmute(data1, j_p = factor(ifelse(j > p, 'judging', 'perceiving')))), aes(x = e_i, fill = j_p)) + 
  geom_bar(position = 'fill', color = 'black') +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = NULL, y = 'Proportion of Participants', fill = NULL)
```

```{r pain-plots}
p1 <- 
  ggplot(data = data1, aes(x = posture, y = pain_1, fill = posture)) +
  geom_boxplot(color = 'black') +
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3, fill = "white") +
  scale_fill_brewer(palette = 'Pastel1') +
  labs(x = 'Posture Type', y = 'Level of Neck Pain') +
  guides(fill = FALSE)

p2 <-
  ggplot(data = data1, aes(x = posture, y = pain_2, fill = posture)) +
  geom_boxplot(color = 'black') +
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3, fill = "white") +
  scale_fill_brewer(palette = 'Pastel1') +
  labs(x = 'Posture Type', y = 'Level of Thoracic Pain') +
  guides(fill = FALSE)

p3 <- 
  ggplot(data = data1, aes(x = posture, y = pain_3, fill = posture)) +
  geom_boxplot(color = 'black') +
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3, fill = "white") +
  scale_fill_brewer(palette = 'Pastel1') +
  labs(x = 'Posture Type', y = 'Level of Lumbar Pain') +
  guides(fill = FALSE)

p4 <-
  ggplot(data = data1, aes(x = posture, y = pain_4, fill = posture)) +
  geom_boxplot(color = 'black') +
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3, fill = "white") +
  scale_fill_brewer(palette = 'Pastel1') +
  labs(x = 'Posture Type', y = 'Level of Sacral Pain') +
  guides(fill = FALSE)

p1 + p2 + p3 + p4
```

# Model Setup

## Data Split & *k*-Fold Cross Validation

```{r}
person_split <- initial_split(data = data1, prop = 0.70, strata = e_i) # initial split

person_train <- training(person_split) # training set
person_test <- testing(person_split) # testing set

person_folds <- vfold_cv(person_train, v = 10, strata = e_i) # 10-fold cross validation
```

## Recipe Building & Workflow

```{r}
simple_rec <-
  recipe(e_i ~ ., data = person_train)

corr_rec <- 
  simple_rec %>%
  step_dummy(all_nominal_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_corr(all_numeric_predictors(),
            threshold = tune())
```

## Model Specifications

### First Recipe (`simple_rec`)

#### Decision Tree
```{r}
tree_spec <- decision_tree(cost_complexity = tune()) %>% 
  set_mode('classification') %>% 
  set_engine('rpart')
```

#### Random Forest
```{r}
forest_spec <- rand_forest(mtry = tune()) %>% 
  set_mode('classification') %>% 
  set_engine('ranger')
```

### Second Recipe (`corr_rec`)

#### K-Nearest Neighbors
```{r}
knn_spec <- nearest_neighbor(neighbors = tune()) %>% 
  set_mode('classification') %>% 
  set_engine('kknn')
```

#### Logistic Regression
```{r}
log_spec <- logistic_reg() %>% 
  set_mode('classification') %>% 
  set_engine('glm')
```

# Model Tuning & Evaluation

## Workflow Sets
```{r}
simple_wf <- 
  workflow_set(
    preproc = list(simple = simple_rec),
    models = list(class_tree = tree_spec,
                  random_forest = forest_spec)
  )

corr_wf <- 
  workflow_set(
    preproc = list(corr = corr_rec),
    models = list(knn = knn_spec,
                  log_reg = log_spec)
  )
```

## Workflow Map & Tuning
```{r eval = FALSE}
all_workflows <-
  bind_rows(simple_wf, corr_wf)

grid_ctrl <-
   control_grid(
      save_pred = TRUE,
      parallel_over = "everything",
      save_workflow = TRUE
   )

grid_results <- all_workflows %>%
   workflow_map(
      seed = 123,
      resamples = person_folds,
      grid = 10,
      control = grid_ctrl
   )

grid_results %>% 
  write_rds('C:/Users/giova/Desktop/PSTAT 131/predicting_personality/predicting_personality_files/Models/tuned_grid.rds')
```

## Evaluation of Models
```{r}
autoplot(
   grid_results,
   rank_metric = "roc_auc",
   metric = "roc_auc",
   select_best = TRUE
)

grid_results %>% 
   rank_results() %>% 
   filter(.metric == "roc_auc") %>% 
   select(model, .config, roc_auc = mean, rank)
```


